 %\documentclass[preprint,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%%  \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%\documentclass[final,5p,times]{elsarticle}
\documentclass[final,5p,times,twocolumn]{elsarticle}

\newcommand{\note}[2][inline]{\todo[color=red!30,#1]{\small\sf #2}}

%% if you use PostScript figures in your article
%% use the graphics package for simple commands
%% \usepackage{graphics}
%% or use the graphicx package for more complicated commands
%% \usepackage{graphicx}
%% or use the epsfig package if you prefer to use the old commands
%% \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{rotating,todonotes, xspace}
\usepackage{amssymb}
\usepackage[section]{placeins}
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{float}
\usepackage{subcaption}
\captionsetup{compatibility=false}
\usepackage{wrapfig}
\usepackage{array}
\usepackage{multicol}
\usepackage{hyperref}
\usepackage{enumitem}

% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}


\journal{Future Generation of Computer Systems}

\begin{document}

\begin{frontmatter}

%Should change the tilte to something better
\title{Using Imbalance Metrics to Optimize Task Clustering in Scientific Workflow Executions}


\author[isi]{Weiwei Chen\corref{cor1}}
\ead{weiweich@acm.org}

\author[isi]{Rafael Ferreira da Silva}
\ead{rafsilva@isi.edu}

\author[isi]{Ewa Deelman}
\ead{deelman@isi.edu}

\author[man]{Rizos Sakellariou}
\ead{rizos@cs.man.ac.uk}

\cortext[cor1]{Corresponding address: USC Information Sciences Institute, 4676 Admiralty Way Ste 1001, Marina del Rey, CA, USA, 90292, Tel: +1 310 448-8408}


\address[isi]{University of Southern California, Information Sciences Institute, Marina del Rey, CA, USA}
\address[man]{University of Manchester, School of Computer Science, Manchester, U.K.}


\begin{abstract}
Scientific workflows can be composed of many fine computational granularity tasks. The runtime of these tasks may be shorter than the duration of system overheads, for example, when using multiple resources of a cloud infrastructure. Task clustering is a runtime optimization technique that merges multiple short running tasks into a single job such that the scheduling overhead is reduced and the overall runtime performance is improved. However, existing task clustering strategies only provide a coarse-grained approach that relies on an over-simplified workflow model. In this work, we examine the reasons that cause Runtime Imbalance and Dependency Imbalance in task clustering. Then, we propose quantitative metrics to evaluate the severity of the two imbalance problems. Furthermore, we propose a series of task balancing methods (horizontal and vertical) to address the load balance problem when performing task clustering for 5 widely used scientific workflows. Finally, we analyze the relationship between these metric values and the performance of these task balancing methods. A trace-based simulation shows our methods can significantly improve the runtime performance of workflow applications when compared to a baseline execution. We also compare the performance of our methods with two algorithms from the literature.

\end{abstract}

\begin{keyword}
Scientific workflows \sep Performance analysis \sep Scheduling \sep Workflow simulation \sep Task clustering \sep Load balancing
\end{keyword}

\end{frontmatter}

\input{introduction.tex}
\input{relatedwork.tex}
\input{model.tex}
\input{imbalance.tex}
\input{evaluation.tex}
\input{futurework.tex}





%% References with bibTeX database:

\bibliographystyle{elsarticle-num}
\bibliography{biblio}

\end{document}