%\documentclass[preprint,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%%  \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%\documentclass[final,5p,times]{elsarticle}
\documentclass[final,5p,times,twocolumn]{elsarticle}

%% if you use PostScript figures in your article
%% use the graphics package for simple commands
%% \usepackage{graphics}
%% or use the graphicx package for more complicated commands
%% \usepackage{graphicx}
%% or use the epsfig package if you prefer to use the old commands
%% \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}

\usepackage{makeidx}  % allows for indexgeneration
\usepackage{algorithm,algpseudocode}
\usepackage{graphicx}
\usepackage{float}
\usepackage{subcaption}
\captionsetup{compatibility=false}
\usepackage{wrapfig}
\usepackage{array}
\usepackage{multicol}

%\usepackage{cite}
% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}


\journal{Future Generation of Computer Systems}

\begin{document}

\begin{frontmatter}


\title{Scientific Workflow Analysis with Quantitative and Structural Metrics}


\author[isi]{Weiwei Chen\corref{cor1}}
\ead{weiweich@acm.org}

\author[isi]{Rafael Ferreira da Silva}
\ead{rafsilva@isi.edu}

\author[isi]{Ewa Deelman}
\ead{deelman@isi.edu}


\author[man]{Rizos Sakellariou}
\ead{rizos@cs.man.ac.uk}

\cortext[cor1]{Corresponding address: USC Information Sciences Institute, 4676 Admiralty Way Ste 1001, Marina del Rey, CA, USA, 90292, Tel: +1 310 448-8408}


\address[isi]{University of Southern California, Information Sciences Institute, Marina del Rey, CA, USA}
\address[man]{University of Manchester, School of Computer Science, Manchester, U.K.}


\begin{abstract}
As scientific workflows grow in complexity and importance, the designers of scientific workflows need a deeper and broader understanding of the characteristics and features of workflows and how they behave in order to improve the overall performance. 
This paper provides a quantitative, structural, and overhead aware analysis of workflows of diverse scientific applications, including astronomy, bioinformatics, earthquake science, and gravitational-wave physics. The workflow analysis is based on novel workflow structural metrics that provide substantial information about the characteristics of workflow structures and concretely show how they influence the overall performance. 
%This information includes density, overlap, and sensitivity etc.. 
This paper further applies these metrics to three popular workflow research scenarios including workflow profiling, task clustering, and overhead aware task scheduling. Trace based analysis and simulations show their strong capabilities in revealing hidden information and driving improvements on workflow performance. 

\end{abstract}

\begin{keyword}
Scientific workflows \sep Performance analysis \sep Scheduling \sep Workflow Simulation
\end{keyword}

\end{frontmatter}

\input{introduction.tex}
\input{model.tex}
\input{profiling.tex}
\input{imbalance.tex}
\input{sensitivity.tex}

\section{Conclusion, Discussion and Future Work}

In this work, we introduced a series of quantitative and structural metrics in revealing workflow structure information and we further applied them to three usage scenario to show their influence with traced based analysis and simulations. The experiments show that these metrics can significantly improve the performance of task clustering strategies and task scheduling strategies.   

In our work, we have shown that these metrics have strong connection with the performance of different optimization methods. However, not all of them perform well under all of the circumstances. Also, the origin of these metrics are heavily depending on human intuition. Still, there is a lack of understanding of whether we can have more and better metrics in revealing how workflow structures interact with different optimization methods. As an initial effort, we will continue to use pattern discovery techniques to collect as many metrics as possible and performance experiments on them. 

With enough metrics, one important step we are going to take is to aggregate multiple metrics and use machine learning techniques to reveal their correlation. However, combining the metrics we have proposed above is challenging. First, these metrics are quantitative and can be measured, however, on their own they are hard to compare against each other and determine whether one metric is more significant than others. Second,  having multiple metrics together increase the complexity and this is against our initial goal, which is reduce the complexity in analyzing workflows. A more sophisticated alternative, however, is to conduct a statistical analysis based on Principal Component Analysis (PCA). This method determines in a more precise way the existing correlation between the collected metrics and their impact on workflow performance.

%A more sophisticated alternative, however, is to conduct a statistical analysis based on Principal Component Analysis (PCA). This method determines in a more precise way the existing correlation between the collected metrics and their impact on resilience. However, undertaking an automated analysis using PCA requires prior experimental data to be available. For this reason, in or- der to better enable a comparison, the aggregation of all the metrics into a single value is proposed in this section. Aggregations of multiple variables are used in many well- known disciplines such as statistical or economic measure- ment. However, combining the metrics we have proposed above provides three main challenges. First, the metrics are quantitative and can be measured, however on their own they are hard to compare against each other and to deter- mine whether one metric is more significant than another. Second, having multiple metrics makes any comparison difficult. Third, resilience increases in direct proportion to some metrics, whereas it reduces with respect to others. Identify- ing to what degree each measured metric impacts resilience (directly or inversely) must also be qualified through previously recorded experimental data or by an expert.

%% References with bibTeX database:

\bibliographystyle{elsarticle-num}
\bibliography{biblio}

%% Authors are advised to submit their bibtex database files. They are
%% requested to list a bibtex style file in the manuscript if they do
%% not want to use elsarticle-num.bst.

%% References without bibTeX database:

% \begin{thebibliography}{00}

%% \bibitem must have the following form:
%%   \bibitem{key}...
%%

% \bibitem{}

% \end{thebibliography}


\end{document}

%%
%% End of file `elsarticle-template-num.tex'.
